NURM: a tool to locate numerical differences in pipelines

### 2019-05-27 paper.tex
Edits and comments on subject types

\usepackage{subcaption}
\usepackage{xspace}

@@

We use \reprozip to record: (1) 
the tree of processes executed by the pipeline and (2) the list of 
files read and written by each process. This information is collected 
by \reprozip through the \texttt{ptrace} Unix system 
call, and stored in a \texttt{SQLite} database. The database stores 
information of all the processes which are created by the 
\texttt{clone()} or \texttt{fork()} system calls. It also records the 
files opened by each process (in read, write or execution mode), 
including result files and temporary files.

Our tool reconstructs the tree of processes starting from the first process
created by the pipeline and traversing the tree through occurrences of
\texttt{clone()} and \texttt{fork()}. It also creates a process
\emph{graph} from this tree by adding edges corresponding to file
dependencies between processes. A file dependency is defined between
processes A and B if a file written by A 
is read by B. Figure~\ref{fig:simple_script} shows an example of a 
process tree and graph constructed from the example pipeline in
Listing~\ref{listing:sample-script}.
\begin{listing}
\begin{minted}[frame=single,
  framesep=3mm,
  linenos=false,
  xleftmargin=0pt,
  tabsize=4]{bash}
#!/bin/bash
if [ #$ !=1 ]
then
    echo "usage: 0 <inputimage.nii.gz>"
    exit 1
fi

@@

Different subjects may lead to different process trees and graphs, due to
variations in the processing pipeline. Such variations may be coming from differences
in data types or cardinality, for instance, some subj

Depending on the variety of computing environments, in particular, 
different analysis software packages, different process trees can be 
constructed for the same experiment. It was expected to achieve the 
same process trees since we used the same computing environments 
specially the same analysis software package in different conditions. 
However, we found that different subjects may result in different 
process trees.

Therefore, all subjects should be grouped into different types based on 
the topological structure of the process trees.
To this order, the distance between trees, topological differences, 
need to be calculated.

We used the tree edit distance~\cite{zhang1989simple} method to measure 
the distance between labeled trees. The distance between two labeled 
trees is defined as the minimum weighted number of edit operations that 
it takes to transform one tree to another. There are three edit 
operations that can be used to transform a tree including modify, 
remove, and insert. Modifying node \textit{n} means changing the label 
on \textit{n}. Removing node \textit{n} means making the children of 
\textit{n} become the children of its parent and then 
deleting it. Inserting is the complement of remove; this 
means that inserting \textit{n} as the child of \textit{n}$'$ will make 
\textit{n} the parent of a consecutive subsequence of the current 
children of \textit{n}$'$.
Each operation has an associated cost---insertions and removals each 
cost 1, and modifications cost either 1 or 2, depending on whether 
you consider a modification to essentially be just a removal 
or followed by an insertion.

We clustered subjects into various types using the bottom-up hierarchical 
clustering approach implemented in \textit{scipy} python package (see 
Algorithm~\ref{algo:hclustering}). This approach groups together the 
subjects with similar distance, less than a threshold.

% \texttt{agglomerative approach}
%  based on \texttt{nearest\_neighbors} method

\begin{algorithm}[h!]
\caption{Hierarchical clustering algorithm from \textit{scipy}}
\label{algo:hclustering}
\begin{algorithmic}

  \State /* Input: a list of n trees; \texttt{T = [t${_1}$, t${_2}$,...,t${_n}$]}
  \State /* Output: a list of clusters; \texttt{C = [c${_1}$,c${_2}$,...,c${_k}$]}
  \For{\texttt{i=1} in \texttt{n}}
  \State \texttt{C${_i}$} = \texttt{T${_i}$} 
  \State \texttt{\# Create a cluster for each tree in input list;}
  \EndFor
  \While{\texttt{len(C)} > \texttt{1}}  
  \For{\texttt{i=1} to \texttt{n}}
  \For{\texttt{j=i+1} to \texttt{n}}
  \State \texttt{D${_{ij}}$} = \texttt{edit\_distance(C${_i}$, C${_j}$)}
  \State \texttt{\# Select two nearest clusters;}
  \EndFor
  \EndFor
  \State \texttt{d\_min${_{ij}}$} = \texttt{MIN(D)}
  \If{\texttt{d\_min${_{ij}}$} > \texttt{thre}}
  \State Return \texttt{C}
  \Else {
  \State \texttt{C${_i}$} = Merge(\texttt{C${_i}$ and C${_j}$})
  \State Remove \texttt{C${_j}$}}
  \EndIf
  \EndWhile
%  \State Function edit\_distance (C${_1}$, C${_2}$)
%  \State EndFunction
\end{algorithmic}
\end{algorithm}

The clustering algorithm starts by finding two closest subjects to each 
other on the basis of edit distance method and form a new cluster 
between them. In the next iteration, this cluster aggregates with the 
next nearest cluster or subject and turns to another cluster. This 
process continues until all the subjects are aggregated to form one big 
cluster or the distances are less than an specified threshold. The 
threshold is defined as the minimum distance required to separate 
clusters.

>>>>

\usepackage{subcaption}
\usepackage{xspace}
\usepackage{stmaryrd} % for llbracket and rrbracket
\usepackage{amsmath}
\DeclareMathOperator*{\argmin}{argmin}

@@

We use \reprozip to record: (1) the set of processes created by the
pipeline, using the \texttt{clone()} or \texttt{fork()} system call, and
(2) the set of files read, written and executed by each process, including
temporary files. \reprozip collects this information through the
\texttt{ptrace()} system call and stores it in a SQLite database.

Our tool reconstructs a process tree starting from the first process
created by the pipeline and traversing processes through \texttt{clone()}
and \texttt{fork()}. It also creates a process \emph{graph} from this tree
by adding edges corresponding to file dependencies between processes. A
file dependency is defined between processes A and B if a file written by A
is read by B. Figure~\ref{fig:simple_script} shows an example of a process
tree and graph constructed from the example pipeline in
Listing~\ref{listing:sample-script}.
\begin{listing}
\begin{minted}[frame=single,
  framesep=3mm,
  linenos=false,
  xleftmargin=0pt,
  tabsize=4]{bash}
#!/bin/bash
if [ #$ !=1 ]
then
    echo "usage: 0 <inputimage.nii.gz>"
    exit 1
fi

@@

Different subjects may lead to different process trees and graphs, due to
variations that may be coming from differences in data types or
cardinality. For instance, some images may have been acquired multiple
times in some subjects. Some of these differences can be neglected, for
instance when a data decompression step is present at the beginning of the
execution for some subjects only, and other ones cannot, when different
processing paths are used in different subjects.

To address this issue, we use the tree edit distance~\cite{zhang1989simple}
to measure the distance between labeled trees \tristan{how about graphs?
also, why between \emph{labeled} trees and not just trees?}, defined as the
minimum weighted \tristan{weighted by what?} number of edit operations
required to transform one tree into the other one. Three edit operations
are considered: node label modification, node removal, and node insertion.
Each operation has an associated cost: 1 for insertion and removal, and
either 1 or 2 for label modification depending on the type of modification. \tristan{Please check my edits.}

With this distance, we cluster subjects using agglomerative hierarchical
clustering as implemented in SciPy~\cite{oliphant2007scipy}
(Algorithm~\ref{algo:hclustering}). As a result, the distance between
any two subjects in a cluster is lower than a threshold parameter.

\begin{algorithm}[h!]
\caption{Hierarchical clustering algorithm from SciPy}
\label{algo:hclustering}
\begin{algorithmic}

  \State /* Input: - a list of n trees; \texttt{T = [t${_1}$, t${_2}$,...,t${_n}$]}
  \State /*\quad \quad \quad \quad - a distance threshold value; \texttt{threshold}
  \State /* Output: a list of clusters; \texttt{C = [c${_1}$,c${_2}$,...,c${_k}$]}
  \State \texttt{\# Create a cluster for each tree in input list;}
  \State \texttt{C = [t${_1}$, t${_2}$,...,t${_n}$]}
  \State \texttt{\# Define a 2 $\times$ 2 array to store distances}
  \State \texttt{D} = \texttt{array(n, n)}
  \While{\texttt{len(C)} > \texttt{1}}  
  \State \texttt{\# Select the two nearest clusters}
  \For{\texttt{i=1} to \texttt{n}}
  \For{\texttt{j=i+1} to \texttt{len(C)}}
  \State \texttt{D[i][j]} = \texttt{edit\_distance(C[i], C[j])}
    \EndFor
  \EndFor
  \State \texttt{i, j} = $\argmin \left\{ \texttt{D[i][j]}, \ \texttt{i, j} \in \llbracket 1, \texttt{len(C)}\rrbracket^2 \right\}$
  \If{\texttt{D[i][j]} > \texttt{threshold}}
  \State \Return \texttt{C}
  \Else
  \State \texttt{\# Merge C[j] into C[i]}
  \State \texttt{C[i]} = \texttt{merge}(\texttt{C[i] and C[j]})
  \State \texttt{\# Remove C[j] from C}
  \State \texttt{remove(C[j], C)}
  \EndIf
  \EndWhile
%  \State Function edit\_distance (C${_1}$, C${_2}$)
%  \State EndFunction
\end{algorithmic}
\end{algorithm}
